\documentclass[letterpaper,11pt]{article}
\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
\usepackage{settings_custom}
\addbibresource{refs.bib}
\allowdisplaybreaks


\begin{document}
\title{Average-case matrix discrepancy: satisfiability bounds}
\date{}
\author{}

\maketitle

\todo{
\begin{enumerate}
    \item Fix the anonymity, I need to re-read to make sure I don't disclose my identity. \done
    \item Finish writing the proof for the failure of second moment.
   \item Add a sentence that in upcoming work we will characterize the second moment sharply using methods from statistical physics, as 
well as the typical number of solutions.
    \item Re-read all new passages, and the new Section and proof. 
    \item Check that there is no missing mention of the new results in the rest.
\end{enumerate}
}

\vspace{-26pt}

\begin{abstract}
    Given a sequence of $d \times d$ symmetric matrices $\{\bW_i\}_{i=1}^n$, and a margin $\Delta > 0$, we investigate whether it is possible to find signs $(\eps_1, \dots, \eps_n) \in \{\pm 1\}^n$ such that the operator norm of the signed sum satisfies $\|\sum_{i=1}^n \eps_i \bW_i\|_\op \leq \Delta$.
    \cite{kunisky2023online} recently introduced a random version of this problem, where the matrices $\{\bW_i\}_{i=1}^n$ are drawn from the Gaussian orthogonal ensemble. 
    This model can be seen as a random variant of the celebrated Matrix Spencer conjecture, and as a matrix-valued analog of the symmetric binary perceptron in statistical physics.
%
    In this work, we establish a satisfiability transition in this problem as $n, d \to \infty$ with $n / d^2 \to \tau > 0$.
    First, we prove that the expected number of solutions with margin $\Delta = \kappa \sqrt{n}$ has a sharp threshold at a critical $\tau_1(\kappa)$: 
    for $\tau < \tau_1(\kappa)$ the problem is typically unsatisfiable, while for $\tau > \tau_1(\kappa)$ the average number of solutions becomes exponentially large.
    Second, combining a second-moment method with recent results from \cite{altschuler2023zero} on margin concentration in perceptron-type problems, we identify a second threshold $\tau_2(\kappa)$, 
    such that for $\tau > \tau_2(\kappa)$ the problem admits solutions with high probability.
%
    In particular, we establish for the first time that a system of $n = \Theta(d^2)$ Gaussian random matrices can be balanced so that the spectrum of the resulting matrix macroscopically shrinks compared to the typical semicircle law.
    \new{
    Finally, we prove that there exists values of $(\tau,\kappa)$ for which the average number of solutions is both exponentially large and does not concentrate around its average, 
    implying the failure of the second moment method, and uncovering a richer picture than in the symmetric binary perceptron.}
    Our proofs rely on deriving concentration inequalities and large deviation properties for the law of correlated Gaussian matrices under spectral norm constraints, extending results in the literature and offering insights of independent interest.
\end{abstract}

\setcounter{tocdepth}{2}
\tableofcontents

\section{Introduction and main results}\label{sec:intro}
\input{sections/introduction.tex}

\section{Sharp asymptotics of the first moment}\label{sec:1st_moment}
\input{sections/first_moment.tex}

\section{The satisfiability region}\label{sec:2nd_moment}
\input{sections/second_moment.tex}

\section{Limitations of the second moment method}\label{sec:fail_2nd_moment}
\input{sections/fail_2nd_moment.tex}

\printbibliography
\end{document}
