\documentclass[a4paper,11pt]{article}
\usepackage[top=2.5cm, bottom=2.5cm, left=2.5cm, right=2.5cm]{geometry}
\usepackage{settings_custom}
\addbibresource{refs.bib}
\allowdisplaybreaks

\newcommand{\changenew}[1]{{\color{Orchid} #1}}


\begin{document}
\title{Average-case matrix discrepancy: satisfiability bounds}
\date{}
\author{Antoine Maillard\thanks{\texttt{antoine.maillard@inria.fr}.}}
\affil{Inria Paris \& DIENS, Ecole Normale Sup\'erieure, PSL Research University. 
\\ Department of Mathematics, ETH Z\"urich, Switzerland.}

\maketitle

\begin{abstract}
    Given a sequence of $d \times d$ symmetric matrices $\{\bW_i\}_{i=1}^n$, and a margin $\Delta > 0$, we investigate whether it is possible to find signs $(\eps_1, \dots, \eps_n) \in \{\pm 1\}^n$ such that the operator norm of the signed sum satisfies $\|\sum_{i=1}^n \eps_i \bW_i\|_\op \leq \Delta$.
    \cite{kunisky2023online} recently introduced a random version of this problem, where the matrices $\{\bW_i\}_{i=1}^n$ are drawn from the Gaussian orthogonal ensemble. 
    This model can be seen as a random variant of the celebrated Matrix Spencer conjecture, and as a matrix-valued analog of the symmetric binary perceptron in statistical physics.
%
    In this work, we establish a satisfiability transition in this problem as $n, d \to \infty$ with $n / d^2 \to \tau > 0$.
    Our main results are twofold.
    First, we prove that the expected number of solutions with margin $\Delta = \kappa \sqrt{n}$ has a sharp threshold at a critical $\tau_1(\kappa)$: 
    for $\tau < \tau_1(\kappa)$ the problem is typically unsatisfiable, while for $\tau > \tau_1(\kappa)$ the average number of solutions becomes exponentially large.
    Second, combining a second-moment method with recent results from \cite{altschuler2023zero} on margin concentration in perceptron-type problems, we identify a second threshold $\tau_2(\kappa)$, 
    such that for $\tau > \tau_2(\kappa)$ the problem admits solutions with high probability.
%
    In particular, we establish that a system of $n = \Theta(d^2)$ Gaussian random matrices can be balanced so that the spectrum of the resulting matrix macroscopically shrinks compared to the typical semicircle law.
    Finally, under a technical assumption, we show that there exists values of $(\tau,\kappa)$ for which the number of solutions has large variance, 
    implying the failure of the second moment method and uncovering a richer picture than in the vector-analog symmetric binary perceptron problem.
    Our proofs rely on concentration inequalities and large deviation properties for the law of correlated Gaussian matrices under spectral norm constraints.
\end{abstract}

\setcounter{tocdepth}{2}
\tableofcontents

\section{Introduction and main results}\label{sec:intro}
\input{sections/introduction.tex}

\subsection*{Acknowledgements}
The author wants to thank Dylan Altschuler for discussions and a careful reading of a first version of this manuscript, 
as well as Tim Kunisky and Afonso S.\ Bandeira for insightful discussions about the early stages of this work.

\section{Sharp asymptotics of the first moment}\label{sec:1st_moment}
\input{sections/first_moment.tex}

\section{The satisfiability region}\label{sec:2nd_moment}
\input{sections/second_moment.tex}

\section{Limitations of the second moment method}\label{sec:fail_2nd_moment}
\input{sections/fail_2nd_moment.tex}

\appendix 
\section{Large deviations for the spectral norm of a Gaussian matrix}\label{sec:appendix_ldp}
\input{sections/appendix.tex}

\printbibliography
\end{document}
